Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-11 paper/paper.tex
sec 1 done.

The implementation  of our  analysis annotates each  potential garbage
collection  point  in   the  program  with  a   set  of  deterministic
finite-state automata (DFA) describing the  liveness at that point. We
also implement a copying collector that  uses the DFA to preserve live
objects.   Our  experiments  confirm  that  liveness-based  collection
always increases  the garbage  reclaimed, consequently  decreasing the
number of collections.  In addition,  it reduces the execution time of
several programs.  Experiments  also reveal that the  best results are
obtained  for   a  mixed  garbage   collection  scheme  that   does  a
liveness-based    collection   for    fully    evaluated   data    and
reachability-based   collection  for   closures.   On   our  benchmark
programs,  we  obtain a  decrease  of  up to  45\%  in  the number  of
collections, and  a decrease of  up to 88\%  in the number  of dragged
cells.  Even  though  liveness   based  collector  uses  an  expensive
traversal scheme,  the execution  time is  competitive (0.2X  - 2.08X)
when compared to a reachability based collector.
\end{abstract}

@@

{\color  {Myblue}   Experiments  with  a  single   generation  copying
  collector confirm the expected performance benefits.  Liveness-based
  collection  results  in an  increase  in  garbage reclaimed.   As  a
  consequence, there is a decrease in  the number of collections and a
  (some quantification)  decrease in the  memory size required  to run
  programs. Besides there is a reduction in the overall execution time
  for  some  programs.  Our  experiments  further  show that  a  mixed
  strategy, where we  do liveness based GC up to  the closure boundary
  and reachability based GC for  the arguments (free variables) of the
  closure,  gives us  similar benefits  as a  fully liveness  based GC
  (where  the  arguments  of  the closure  are  also  collected  using
  liveness)  at  much  less  overhead (?).  [Overhead:  extra  memory,
    book-keeping during run-time to update, multiple traversal of same
    set of cells][Refer to Section~\ref{some-sec-in-impl}] }

@@

\subsection{Copying closure using liveness vs copying closures using reachability}

>>>>

The implementation  of our  analysis annotates each  potential garbage
collection  point  in   the  program  with  a   set  of  deterministic
finite-state automata (DFA) describing the  liveness at that point. We
also implement a copying collector that  uses the DFA to preserve live
objects.   Our  experiments  confirm  that  liveness-based  collection
always increases  the garbage  reclaimed, consequently  decreasing the
number of collections.  In addition,  it reduces the execution time of
several programs.  Experiments  also reveal that the  best results are
obtained  for   a  mixed  garbage   collection  scheme  that   does  a
liveness-based    collection   for    fully    evaluated   data    and
reachability-based   collection  for   closures.   On our benchmark
programs, we obtain a decrease of up to 45\% in the number of
collections, and a decrease of up to 88\% in the number of dragged
cells.  Even though liveness based collector uses an expensive
traversal scheme, the execution time is competitive (0.2X - 2.1X) when
compared to a reachability based collector.
\end{abstract}

@@

Experiments    with   a    single    generation   copying    collector
(Section~\ref{sec:experiments})  confirm   the  expected  performance
  benefits.   Liveness-based  collection  results in  an  increase  in
  garbage reclaimed.   As a  consequence, there is  a decrease  in the
  number of  collections (up to 45\%)  and a decrease (up  to 88\%) in
  the number  of {\em dragged}  cells (cells  that survive a  GC after
  their  last  use). Besides  there  is  a  reduction in  the  overall
  execution time for some programs.  Our experiments show that a mixed
  strategy, where we  do liveness based GC up to  the closure boundary
  and reachability based GC for  the arguments (free variables) of the
  closure,  gives us  similar benefits  as a  fully liveness  based GC
  (where  the  arguments  of  the closure  are  also  collected  using
  liveness) at much less overheads (Section~\ref{sec:strategies}).

@@

\subsection{Copying closure using liveness vs copying closures using reachability}\label{sec:strategies}
